---
title: "Simulate the data"
author: "A Solomon Kurz"
date: "`r format(Sys.Date())`"

---

## What?

The goal of this file is to simulate non-linear RCT data that have some of the characteristics of the CAPS data in Reger et al (2016; https://dx.doi.org/10.1037/ccp0000134), which you can find posted on Research Gate at https://tinyurl.com/2p88xk2n.

## Load

Load the necessary R packages.

```{r, warning = F, message = F}
library(tidyverse)
library(faux)
```

## Simulate

In the study, you can find summary statistics for the CAPS for three conditions. Here, we'll jsut simulate the data for the two treatment conditions:

* prolonged exposure (PE) and
* virtual reality exposure therapy (VRET).

For simplicity, I'm just assuming the test-retest reliability is .5 among all time points (which is probably pretty unlikely).

Simulate the data for the PE condition.

```{r}
# initial sample size per group
n <- 54

# simulate MSEL and VABS values for the A/M group
set.seed(1)

pe <- rnorm_multi(
  n = n,
  mu = c(78.28, 65.03, 44.28, 36.63, 38.33),
  sd = c(16.35, 29.19, 33.73, 31.80, 28.49), 
  r = .5, 
  varnames = list("caps.0", "caps.1", "caps.2", "caps.3", "caps.4")
)
```

Simulate the data for the VRET condition.

```{r}
# initial sample size per group
n <- 54

# simulate MSEL and VABS values for the A/M group
set.seed(2)

vret <- rnorm_multi(
  n = n,
  mu = c(80.44, 71.19, 57.07, 56.64, 53.50),
  sd = c(16.23, 23.27, 32.32, 31.50, 28.07), 
  r = .5, 
  varnames = list("caps.0", "caps.1", "caps.2", "caps.3", "caps.4")
)
```

Join the two data sets and wrangle.

```{r}
d <- bind_rows(pe, vret) %>% 
  mutate(id = 1:n(),
         tx = rep(c("pe", "vret"), each = n() / 2)) %>% 
  pivot_longer(contains("caps"), values_to = "caps") %>% 
  mutate(wave = str_remove(name, "caps.") %>% as.double()) %>% 
  mutate(caps = ifelse(caps < 0, 0,
                       ifelse(caps > 123, 123, caps))) %>% 
  mutate(weeks = case_when(
    wave == 0 ~ 0,
    wave == 1 ~ 2.5,
    wave == 2 ~ 5,
    wave == 3 ~ 12,
    wave == 4 ~ 26
  )) %>% 
  mutate(group = ifelse(tx == "pe", "a", "b"))

d
```

## EDA

Explore the data in a couple plots.

```{r}
d %>% 
  ggplot(aes(x = weeks)) +
  geom_line(aes(y = caps, group = id),
            size = 1/6, alpha = 1/2) +
  geom_line(data = . %>% 
              group_by(weeks, group) %>% 
              summarise(caps = mean(caps)),
            aes(y = caps),
            size = 1.5) +
  geom_linerange(data = . %>% 
                   group_by(weeks, group) %>% 
                   summarise(m = mean(caps),
                             s = sd(caps)),
                 aes(y = m, ymin = m - s, ymax = m + s)
                 ) +
  scale_x_continuous(breaks = c(0, 2.5, 5, 12, 26),
                     labels = c("0", "2.5", "5", "12", "26")) +
  labs(subtitle = "Dark lines are the group means and the vertical lines are plus/minus one standard deviation.\nThe thinner lines are the participant-level trajectories.",
       y = "dv") +
  facet_wrap(~ group, labeller = label_both) +
  theme(panel.grid = element_blank())
```

```{r}
d %>% 
  mutate(wave = factor(wave,))
  ggplot(aes(x = wave, y = caps)) +
  geom_boxplot() +
  facet_wrap(~ tx)
```

## Session information

```{r}
sessionInfo()
```







